Binary Latent Representations for Efficient Ranking

### 2017-08-14 paper/paper.tex
Fix typos.

\section{Introduction}
Industry ranking and recommendation systems need to scale to tens or hundreds of millions of items and users. As the number of items available to be recommended grows large, efficiently ranking the item catalogue to produce top-ranked recommendations becomes increasingly challenging. With large catalogues, ranking latency (in on-line settings) and compute and storage costs (in pre-compute settings) place significant contraints on the design of the entire recommender system.

@@

Additionaly, while a pre-compute solution removes model prediction from the critical path, it still requires substantial investment in computational and storage resources to calculate and cache the predictions of the model. Often the amount of comutation required makes timely updating of the cached results impossible. At Pinterest, pre-calculation of related pins results was so demanding that ranking of different segments of the catalogue had to be staggered in time, leading to stale results and reduced ease of development \citep[Section 5.5]{liu2017related}. \citet{koenigstein2012efficient} estimate that computing recommendations for the Yahoo! Music dataset \citep{dror2011yahoo} with a 50-dimensional latent factor model woud take over 135 hours.

@@

To quantify the tradeoffs involved, we construct and fit both a real-valued and a binary version of simple implicit learning-to-rank factorization model. We introduce the notation and the models below.

@@

As expected, for the same dimensionality a binary model achieves lower ranking accuracy. On average, the accuracy loss when moving from a contious to a binary model of the same latent dimensionality is around 11\%, varying between 6\% and 23\%.

Unfortunately, while continous models retain good accuracy as latent dimensionality decreases, binary models' representational power sharply deteriorates. Moving from the 1024 to 32 dimensions in the continuous model implies a 29 times increase in prediction speed at the expense of a modest 4\% decrease in accuracy. This compares favourably to switching to binary representations: moving to a 1024-binary dimensional representation implies a sharper accuracy drop at 6\% in exchange for a smaller 20 times increase in prediction speed. Moving to 32 dimensions yields a further speed gains at 86 times, but at the cost of a considerable loss of accuracy at 26\%.

>>>>

\section{Introduction}
Industry ranking and recommendation systems need to scale to tens or hundreds of millions of items and users. As the number of items available to be recommended grows large, efficiently ranking the item catalogue to produce top-ranked recommendations becomes increasingly challenging. With large catalogues, ranking latency (in on-line settings) and compute and storage costs (in pre-compute settings) place significant constraints on the design of the entire recommender system.

@@

Additionally, while a pre-compute solution removes model prediction from the critical path, it still requires substantial investment in computational and storage resources to calculate and cache the predictions of the model. Often the amount of computation required makes timely updating of the cached results impossible. At Pinterest, pre-calculation of related pins results was so demanding that ranking of different segments of the catalogue had to be staggered in time, leading to stale results and reduced ease of development \citep[Section 5.5]{liu2017related}. \citet{koenigstein2012efficient} estimate that computing recommendations for the Yahoo! Music dataset \citep{dror2011yahoo} with a 50-dimensional latent factor model woud take over 135 hours.

@@

To quantify the trade-offs involved, we construct and fit both a real-valued and a binary version of simple implicit learning-to-rank factorization model. We introduce the notation and the models below.

@@

As expected, for the same dimensionality a binary model achieves lower ranking accuracy. On average, the accuracy loss when moving from a continuous to a binary model of the same latent dimensionality is around 11\%, varying between 6\% and 23\%.

Unfortunately, while continuous models retain good accuracy as latent dimensionality decreases, binary models' representational power sharply deteriorates. Moving from the 1024 to 32 dimensions in the continuous model implies a 29 times increase in prediction speed at the expense of a modest 4\% decrease in accuracy. This compares favourably to switching to binary representations: moving to a 1024-binary dimensional representation implies a sharper accuracy drop at 6\% in exchange for a smaller 20 times increase in prediction speed. Moving to 32 dimensions yields a further speed gains at 86 times, but at the cost of a considerable loss of accuracy at 26\%.
